---
title: "P8105_hw2_tg2753"
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
```


## Problem 1
Read and clean the Mr. Trash Wheel sheet:  
-specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
-use reasonable variable names  
-omit rows that do not include dumpster-specific data  
-round the number of sports balls to the nearest integer  
```{r}
trashwheel_df = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "Mr. Trash Wheel",
        range = cell_cols("A:N")) %>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(sports_balls = round(sports_balls))
```

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable.
```{r}
precipitation_df_2018 = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "2018 Precipitation",
        skip = 1) %>% 
    janitor::clean_names() %>% 
    drop_na(month) %>% 
    mutate(year = 2018, month = month.name) 

precipitation_df_2019 = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "2019 Precipitation",
        skip = 1) %>% 
    janitor::clean_names() %>% 
    drop_na(month) %>% 
    mutate(year = 2019, month = month.name)

precipitation_df = left_join(precipitation_df_2018, precipitation_df_2019,
                             by = "month")
```
trashwheel_df has `r nrow(trashwheel_df) ` observations while the precipitation_df has `r nrow(precipitation_df)` observations.   
the total precipitation in 2018 is `r sum(pull(precipitation_df, total.x))`  
the median number of sports balls in a dumpster in 2019 is `r median(pull(trashwheel_df,sports_balls))`

## Problem 2
First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r}
pols = read_csv("./Dataset/pols-month.csv") %>% 
    separate(col = mon, into = c("year", "month", "day")) %>% 
    mutate(year = as.character(year),
           month = as.integer(month),
           month = month.name[month],
           president = ifelse(prez_dem=="1","dem","gop")) %>% 
    select(-prez_dem, -prez_gop, -day) 
```

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp = read_csv("./Dataset/snp.csv") %>% 
    mutate(date = lubridate::mdy(date)) %>%
    separate(col = date, into = c("year", "month", "day")) %>% 
    mutate(year = as.numeric(year),
           year = ifelse(year > 2021, year - 100, year)) %>% 
    arrange(year,month) %>%
    mutate(year = as.character(year),
           month = as.integer(month),
           month = month.name[month]) %>%
    select(-day)
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
```{r}
unemployment = read_csv("./Dataset/unemployment.csv") %>% 
     pivot_longer(Jan:Dec,
                 names_to = "month",
                 values_to = "unemployment_percentage") %>%
    mutate(Year = as.character(Year),
           month = match(month,month.abb),
           month = month.name[month]) %>%
    rename(year = Year) %>%
    relocate(year, month) %>% 
    arrange(year,month)
```
Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}
result = left_join(left_join(pols,snp),unemployment)
```

The dataset _**pols**_ has dimension of `r nrow(pols)` x `r ncol(pols)`. It shows the number of national politicians who are democratic or republican at any given time. Then range of years is `r min(range(pull(pols,year)))` - `r max(range(pull(pols,year)))`. The key variables in the _**pols**_ are: `r names(pols)`.

The dataset _**snp**_ has dimension of `r nrow(snp)` x `r ncol(snp)`. It shows the closing values of the S&P stock (Standard & Poor’s) index, which is often used as a representative measure of stock market as a whole. Then range of years is `r min(range(pull(snp,year)))` - `r max(range(pull(snp,year)))`. The key variables in the _**snp**_ are: `r names(snp)`.

The dataset _**unemployment**_ has dimension of `r nrow(unemployment)` x `r ncol(unemployment)`. It shows the percentage of unemployment on the associated date. Then range of years is `r min(range(pull(unemployment,year)))` - `r max(range(pull(unemployment,year)))`. The key variables in the _**unemployment**_ are: `r names(unemployment)`.

The resulting dataset, _**result**_, has dimension  is `r nrow(result)` x `r ncol(result)`. Then range of years is `r min(range(pull(result,year)))` - `r max(range(pull(result,year)))`. The key variables in the _**result**_ are: `r names(result)`.


## Problem 3

