---
title: "P8105_hw2_tg2753"
output: github_document
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1
Read and clean the Mr. Trash Wheel sheet:  
-specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
-use reasonable variable names  
-omit rows that do not include dumpster-specific data  
-round the number of sports balls to the nearest integer  
```{r}
trashwheel_df = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "Mr. Trash Wheel",
        range = cell_cols("A:N")) %>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(sports_balls = round(sports_balls))
```

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable.
```{r}
precipitation_df_2018 = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "2018 Precipitation",
        skip = 1) %>% 
    janitor::clean_names() %>% 
    drop_na(month) %>% 
    mutate(year = 2018, month = month.name) 

precipitation_df_2019 = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "2019 Precipitation",
        skip = 1) %>% 
    janitor::clean_names() %>% 
    drop_na(month) %>% 
    mutate(year = 2019, month = month.name)

precipitation_df = left_join(precipitation_df_2018, precipitation_df_2019,
                             by = "month")
```
trashwheel_df has `r nrow(trashwheel_df) ` observations while the precipitation_df has `r nrow(precipitation_df)` observations.   
the total precipitation in 2018 is `r sum(pull(precipitation_df, total.x))`  
the median number of sports balls in a dumpster in 2019 is `r median(pull(trashwheel_df,sports_balls))`

## Problem 2
First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r message=FALSE, warning=FALSE}
pols = read_csv("./Dataset/pols-month.csv") %>% 
    separate(col = mon, into = c("year", "month", "day")) %>% 
    mutate(year = as.character(year),
           month = as.integer(month),
           month = month.name[month],
           president = ifelse(prez_dem=="1","dem","gop")) %>% 
    select(-prez_dem, -prez_gop, -day) 
```

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r message=FALSE, warning=FALSE}
snp = read_csv("./Dataset/snp.csv") %>% 
    mutate(date = lubridate::mdy(date)) %>%
    separate(col = date, into = c("year", "month", "day")) %>% 
    mutate(year = as.numeric(year),
           year = ifelse(year > 2021, year - 100, year)) %>% 
    arrange(year,month) %>%
    mutate(year = as.character(year),
           month = as.integer(month),
           month = month.name[month]) %>%
    select(-day)
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
```{r message=FALSE, warning=FALSE}
unemployment = read_csv("./Dataset/unemployment.csv") %>% 
     pivot_longer(Jan:Dec,
                 names_to = "month",
                 values_to = "unemployment_percentage") %>%
    mutate(Year = as.character(Year),
           month = match(month,month.abb),
           month = month.name[month]) %>%
    rename(year = Year) %>%
    relocate(year, month) %>% 
    arrange(year,month)
```
Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r message=FALSE, warning=FALSE}
result = left_join(left_join(pols,snp),unemployment)
```
`r knitr::kable(head(result))`
`r knitr::kable(tail(result))`

 **pols** has  `r nrow(pols)` observations. It shows the number of national politicians who are democratic or republican at any given time. Then range of years is `r min(range(pull(pols,year)))` - `r max(range(pull(pols,year)))`. The key variables : `r names(pols[3:9])`.

 **snp** has `r nrow(snp)` observations. It shows the closing values of the S&P stock (Standard & Poor’s) index, which is often used as a representative measure of stock market as a whole. Then range of years is `r min(range(pull(snp,year)))` - `r max(range(pull(snp,year)))`. The key variables are: `r names(snp[3])`.

**unemployment** `r nrow(unemployment)` observations. It shows the percentage of unemployment on the associated date. Then range of years is `r min(range(pull(unemployment,year)))` - `r max(range(pull(unemployment,year)))`. The key variables are  `r names(unemployment[3])`.

**result**, has  `r nrow(result)` observations. The range of years is `r min(range(pull(result,year)))` - `r max(range(pull(result,year)))`. The key variables  are: `r names(result[3:11])`.


## Problem 3
Load and tidy the data
```{r}
baby_name = read_csv("./Dataset/Popular_Baby_Names.csv")%>% 
    janitor::clean_names() %>%
    distinct() %>%
    mutate(
        ethnicity = recode(ethnicity,
                       "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                       "BLACK NON HISP" = "BLACK NON HISPANIC", 
                       "WHITE NON HISP" = "WHITE NON HISPANIC"),
        childs_first_name = str_to_title(childs_first_name)
    )
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.

```{r}
olivia = filter(
    baby_name, childs_first_name == "Olivia" & gender == "FEMALE") %>%
    select(year_of_birth, ethnicity, rank) %>%
    pivot_wider(names_from = "year_of_birth",values_from = "rank")
```

`r knitr::kable(olivia)`

```{r}
top_male_name = filter(baby_name, gender == "MALE" & rank == 1) %>%
    select(year_of_birth, ethnicity, childs_first_name) %>% 
    pivot_wider(names_from = "year_of_birth",
                values_from ="childs_first_name")
```

`r knitr::kable(top_male_name)`

for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).
```{r}
plot_df = filter(baby_name, 
                 gender == "MALE", 
                 ethnicity == "WHITE NON HISPANIC",
                 year_of_birth == 2016)
ggplot(plot_df, aes(x = rank, y = count)) + geom_point()

```

