---
title: "P8105_hw2_tg2753"
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
```


## Problem 1
Read and clean the Mr. Trash Wheel sheet:  
-specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
-use reasonable variable names  
-omit rows that do not include dumpster-specific data  
-round the number of sports balls to the nearest integer  
```{r}
trashwheel_df = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "Mr. Trash Wheel",
        range = cell_cols("A:N")) %>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(sports_balls = round(sports_balls))
```

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable.
```{r}
precipitation_df_2018 = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "2018 Precipitation",
        skip = 1) %>% 
    janitor::clean_names() %>% 
    drop_na(month) %>% 
    mutate(year = 2018, month = month.name) 

precipitation_df_2019 = 
    read_excel("./Dataset/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
        sheet = "2019 Precipitation",
        skip = 1) %>% 
    janitor::clean_names() %>% 
    drop_na(month) %>% 
    mutate(year = 2019, month = month.name)

precipitation_df = left_join(precipitation_df_2018, precipitation_df_2019,
                             by = "month")
```
trashwheel_df has `r nrow(trashwheel_df) ` observations while the precipitation_df has `r nrow(precipitation_df)` observations.   
the total precipitation in 2018 is `r sum(pull(precipitation_df, total.x))`  
the median number of sports balls in a dumpster in 2019 is `r median(pull(trashwheel_df,sports_balls))`

## Problem 2




## Problem 3

